---
title: "Untitled"
author: "Charles Leahan"
date: "11/11/2020"
output: word_document
---

```{r}
library(tidyverse)
library(nycflights13)
```

# Filtering Joins 

Filtering joins match observations in the same way as mutating joins, but affects the observation, not the variables. 

The first table is the primary tables

* semi_join(x,y) - keeps all observations in x that have a match in y 

* anti_join(x,y) - drops all observations in x that have a match in y 


## Semi-joins

Useful for mathcing filtered summary tables back to the original rows 


```{r}
#Top-5 Most Popular Destinations
top5_dest <- flights %>%  
  count(dest,sort = TRUE) %>% 
  head(5)
top5_dest
```


```{r}
#Find each flight that went to one of these destinations 

flights %>% 
  select(month,day,dep_time,arr_time,dep_delay, arr_delay, dest) %>%
  filter(dest %in% top5_dest$dest) %>% 
  head(5)
```

Alternatively, using semi-joins

```{r}
#Chooses only the observations who's dest matches with a value in top5
flights %>% 
  select(month,day,dep_time,arr_time,dep_delay, arr_delay, dest) %>% 
  semi_join(top5_dest,"dest") %>% 
  head(5)
```

## Anti-Joins

Keeps rows that don't have a match.

Useful for diagnosing join mismatches. 

For example, want to find the flights that don't have a matching tail number in the planes table

```{r}
flights %>% 
  anti_join(planes, by = "tailnum") %>% 
  count(tailnum, sort = TRUE)
```


Examples: 

```{r}

#Find the flights whose destination airports are not included in airports

airports0 <- airports %>% 
  rename("dest"="faa")

flights0 <- flights %>% 
  dplyr::select(year:day,dep_time,arr_time,origin,dest) %>% 
  anti_join(airports0, by="dest")

flights0 %>% group_by(dest) %>% summarize(N = n()) %>% arrange(N)

```


# Set Operations 

Expect x and y to have the same variables. 

* intersect(x,y) : returns only observations in both

* union(x,y): return unique observations either or 

* setdiff(x,y): return observations in x, not in y 

## Examples 

```{r}
df1 <- data_frame(x = c(1,2), y = rep(1,2));df1

df2 <- data_frame(x = rep(1,2), y = c(1,2));df2

intersect(df1, df2)
union(df1, df2)

setdiff(df1,df2)
setdiff(df2,df1)

```

# Data Import with readr

__readr functions__

* read_csv() reas comma-delimited files
* read_csv2() reads semicolon-separtated files
* read_tsv() tab-delimited files
* read_delim() reads in files with any delimiter


























```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


