---
title: "Lecture 10"
author: "Charles Leahan"
date: "10/2/2019"
output: word_document
---

Relational Data Manipulation using dplyr

With Two tables there are three families of verbs that work with two tables at a time:

1. Mutating joins - adds new variables to one table from mathcing rows in another
2. Filtering joins - filter observations from one table based on whether or not they mathc an observation
3. Set Operations - comibine the observations in the data sets as if they were elements 

Assumes you have tidy data 

```{r}
library(tidyverse)
library(nycflights13)

```

```{r}
#Inspect Data 

flights_part <- flights %>% 
  dplyr::select(dep_time,arr_time,carrier:dest)
print(flights_part,n=3)

head(flights_part, n=3)
table(flights$origin)
```



keys - unique identifiers (like carrier)

(From airlines) Each airline has a unique carrier code -> (From airports) each aiport has a unique faa code that can be used to connect flights and airports

# Joins

## Mutatating 

* A mutating join allows you to combine variables from two tables.
  
* It first matches observations by their keys, then copies across variables from one table to the other.

* Like mutate(), the join functions add variables to the right.
  
## Inner 

Includes only observations that match in both x and y 

Unmatched rows are not included in the results

```{r}
x <- tibble(
  key = 1:3,
  val_x = c("x1","x2","x3")
)

y <-tibble(
  key = c(1,2,4),
  val_y = c("y1","y2","y3")
)

x %>%  inner_join(y)
```


#Left Joins

Includes all observations of x (regardless of whether they have a match)

Most commonly used because it ensures that you dont lose observations 

Allows for missing values 

```{r}
x %>% left_join(y)
```

## Right Join

Includes all observation in y (regardless of whether they have a match)

```{r}
x %>%  right_join(y)
```


## Full Join

Includes all observations from x and y 

```{r}
x %>% full_join(y)
```



# Duplicate Keys 

While mutating joins are primarily used to add new variables, they can also generate new observations. 

If a match is not unique, a join will add all possible combinations (the cartesian product) of the matching observations 

```{r}
x1 <- tibble(
  key = c(1,1,2,3),
  val_x = c("x1","x2","x3","x4")
)

y1 <-tibble(
  key = c(1,1,2,4),
  val_y = c("y1","y2","y3","y4")
)
```

Note: 

2 x3 y3 is the only unique match

```{r}
x1 %>% left_join(y1)
x1 %>% right_join(y1)

```


# Example: Adding full carrier names to flights

```{r}
flights2 <- flights %>%  
  dplyr::select(year:day,hour,origin,dest,tailnum,carrier)

flights3 <- flights2 %>% 
  left_join(airlines)

print(flights3,n=5)
```


# Controlling how the tables are matched 

1. Null: 

* if the argument by is omitted, dplyr will match by using all variables that appear in both tables

```{r}
#Flights and Weather have year, month, day, hour and orgin in common

flights4 <- flights2 %>% left_join(weather)
flights4
```

2. By = "key", a character. 

* key is a subset of the common variables in two tables 

```{r}
# Flights and Planes both have the variable year, but they mean different things. Want to join by tailnum

flights2 %>%  
  left_join(planes, by = "tailnum")
```

3. A named character vector: by = c("v1" = "v2)

* This will match variable v1 in table x to variable v2 in table y. 

* The variables from x will be used in the ouput.

```{r}
#NAs where destination airport is not present in list of airports

flights5 <- flights2 %>% 
  left_join(airports, c("dest" = "faa"))

print(flights5[c("origin","dest","name")],n=5)
```

```{r}
#NAs where origin airport is not present in list of airports
flights6 <- flights2 %>% 
  left_join(airports, c("origin" = "faa"))

print(flights6[c("origin","dest","name")],n=5)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

