---
title: 'Data Wrangling: One Table'
author: "Charles Leahan"
date: "10/24/2019"
output: word_document
---

```{r, warning=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
library(nycflights13)

```

_Three components of data wrangling_

1. Import Data

2. Tidy Data

3. Transform Data

_Goal_

Get data ready for further analysis - visualization and modeling

_DPLYR_

Grammar of data wrangling

* commonly used data manipulation tools
* fast performance for in memory operations
* abstract the interface between the data manipulation operations and the data source

# Tibbles 

Use $ (name) or [[]] (position) for subsetting

```{r}
class(presidential["name"])
class(presidential[["name"]])
```

Can change default printing behavior

```{r}
print(presidential, n = 2, width = Inf)
```

Can coerce into tiblle 

```{r}
iris_tb <- as_tibble(iris)
```


# Five Important Verbs

* select() selects columns (variables)
* filter() selects rows (observations)
* arrange() reoders the rows 
* mutate() creates new variables with functions of existing variables
* summarize() collapses many values to a single summary

## Select

Select -> selects columns

```{r}
select(presidential, name, party)
```

Options:

```{r, include = FALSE}
#Different options of select(data,options)
select(presidential,1) #first column
select(presidential,start) #"start column"
select(presidential) #Empty
select(presidential,-end) #All but the last
select(presidential,-c(2,3))

#Slicing
flights1<- select(flights, year:arr_time);print(flights1,n=1) #Selects all variables from year to arrival time 
flights2<- select(flights, -(year:day));print(flights2,n=1)   #Selects all the variables except year through day 
```

Helper Functions - select names that match with string given according to some rule

1. contains() 
2. starts_with()
3. ends_with()
4. num_range()

```{r}
flight_time2 <- select(flights, contains("time"))

head(flight_time1)
```

# Filter Rows 

Note:

* Quotation marks used because Republican is a literal value, not a variable name 

```{r}
#Filter only Republican candidates
filter(presidential, party == "Republican")

#Filter years past 2001
filter(presidential, year(start)>=2001) #year from lubridate
```

## Logical Conditions 


Boolean: 

&, | ,! are and or not 

Set Comparison:

x %in% y which is true if x is an element of Y (See third example)

```{r}

#All flights on decemeber 25th

filter(flights,month == 12 & day == 25)

#All flights that had an arrival delay > 2hrs 

filter(flights,arr_delay >=120)
```

```{r}

#All flights that were operated by United, American, or Delta 

unique(flights$carrier) #Find all the carrier codes 

ThreeAirlines <-filter(flights, carrier %in% c("UA","AA","DL"))
table(ThreeAirlines$carrier)
```

## Combining filter and select 

Reads : Take presidential data frame, then filter the Democratic presidents whose start year is grater than 1973, then select the variable name

```{r}
presidential %>%
  filter(start > 1973 & party == "Democratic") %>%
  dplyr::select(name)
```

# Arrange 


```{r}
mypresidents %>% arrange(desc(term_length))
```

Sort by two variables

```{r}
#descending term length and ascending start for each new length 

mypresidents %>% arrange(desc(term_length),start)

#Sort by term length (ascending) and party 

mypresidents <- presidential %>%
  mutate(term_length = (end - start)/dyears(1)) %>% 
  arrange(term_length,party)

mypresidents
```


# Mutate

```{r}
mypresidents <- presidential %>% 
  mutate(term_length = (end-start)/dyears(1))
head(mypresidents,3)
```


## Useful Creation Functions

```{r, include = FALSE}
#log
y <- c(2, 4, 8)
log2(y)

#Exp  
x <- seq(0,1,0.3)
z <- exp(x); z

#Mod
x <- c(10, 5, 3)
x %/% 3
x %% 3

#Cumulative and Rolling aggregates

z <- seq(1,5);z
 
cumsum(z) ; cummean(z) ; cumprod(z) ; cummin(z) ; cummax(z)
```



# Summarize 

Summarize() collapses a data frame a single row 

Every variables in the output is defined by operations performed on vectors - not individual values 

```{r}
mypresidents %>%
  summarize(
    N = n(), 
    first_year = min(year(start)), 
    last_year = max(year(end)),
    num_dems = sum(party == "Democratic"),
    years = sum(term_length),
    avg_term_length = mean(term_length)
  )
```

N:  The function n() simply counts the number of rows (observations), used to ensure data aggregation is done correctly 
first_year:  the smallest start year of the presidents
last_year:  the largest end year of the presidents
num_dems:  the number of Democratic presidents
years:  total term length of the presidents
avg_term_length:  average term length of the presidents


## Use summarize() with group_by()

```{r}

#Compares average term length and sd by party 

mypresidents %>%
  group_by(party) %>% 
  summarize(
    N = n(), 
    avg_term_length = mean(term_length),
    std_term_length = sd(term_length)
  )
```

```{r}

#Compares the median term length 

mypresidents %>% 
  group_by(party) %>% 
  summarize(
    N= n(),
    med_term_length = median(term_length))
```


## Useful Summary Functions 

Counts

n() returns the size of the current group
n_distinct(x) counts the number of distinct values of a variable x
sum(!is.na(x)) counts the number of non-missing values in x

# Handling Missing Values 

To check whether elements of x is NA -> is.na(x)

To remove the missing values prior to computation -> na.rm=TRUE argument in the function.

To filter out all observations with missing values:

<DATA_FRAME> %>% filter(!is.na(<VARIABLE>)

# Extended Example 


What carrier has the longest average departure delay?

```{r}
#Determine Highest Average Departure Delay

delay<- flights %>%    # Create a new dataset
  dplyr::select(year,month,day,dep_delay,carrier) %>%  # select variables of interest
  filter(month == 1 & day == 1 & dep_delay>0) %>%  # pick flights on 1/1/2013 
  group_by(carrier) %>%   # group the data by carrier
  summarize(AvgDelay = mean(dep_delay,na.rm=TRUE)) %>%  # ignore missing values
  arrange(desc(AvgDelay)) # order the summary dataset by average departure delay
delay
```


What destinations have the longest average arrival delay on Dec 25, 2012?

```{r}
# Determines which destinations have the longest average arrival delay on christmas 

ArrDelay<- flights %>%    # Create a new dataset
  dplyr::select(year,month,day,arr_delay,dest) %>%  # select variables of interest
  filter(month == 12 & day == 25 & arr_delay>0) %>%  # pick flights on 1/1/2013 
  group_by(dest) %>%   # group the data by carrier
  summarize(AvgArrDelay = mean(arr_delay,na.rm=TRUE)) %>%  # ignore missing values
  arrange(desc(AvgArrDelay)) # order the summary dataset by average departure delay
print(ArrDelay,n=10)
```


What carrier has the highest average speed to San Francisco?

```{r}
# Which carrier has the highest average speed to San Fran

Speed<- flights %>%    
  dplyr::select(air_time, distance, dest,carrier) %>%  
  filter(dest=="SFO") %>%  # pick flights to SFO 
  mutate(speed=distance/air_time*60,na.rm=TRUE) %>% # calculate speed  
  group_by(carrier) %>%   # group the data by carrier
  summarize(AvgSpeed = mean(speed,na.rm=TRUE)) %>%  # ignore missing values
  arrange(desc(AvgSpeed)) # order the summary dataset by average departure delay
Speed
```

```{r}
# Which Month has the longest departure delay

ldd <- flights %>% 
  dplyr::select(month,dep_delay) %>% 
  filter(dep_delay>0) %>% 
  group_by(month) %>% 
  summarise(
    AvgDptDelay = mean(dep_delay)) %>% 
  arrange(desc(AvgDptDelay))

ldd 

```



































```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

