---
title: "Homework 5"
author: "Charles Leahan"
date: "10/4/2019"
output: word_document
---


# One Table

## Question 1 

Goal: Determine which names are the "newest" - ie, find the distrubtion of median age, by names, for males alive on 1/1/2014. 

__Preliminary__

```{r warning=FALSE}
library(Hmisc)
library(mdsr)
library(babynames)
BabynamesDist <- make_babynames_dist()
head(BabynamesDist)
```

__Preparing the data__

Determine the top 25 "newest" names of - restricted to names given to > 100,000 males.

```{r}
com_male<- BabynamesDist %>%
  filter(sex == "M") %>%
  group_by(name) %>% 
  mutate(N=n()) %>% 
  filter(N>=4) %>% 
  summarise(
    est_num_alive = sum(est_alive_today),
    q1_age = wtd.quantile(age_today, est_alive_today, probs = 0.25),
    median_age = wtd.quantile(age_today, est_alive_today, probs = 0.5),
    q3_age = wtd.quantile(age_today, est_alive_today, probs = 0.75)
  ) %>%
  filter(est_num_alive > 100000) %>% 
  arrange(median_age) %>%
  head(25)

head(com_male)
```

__Creating the mapping and context__

Note: have to re-order based on factor levels, since sort() and arrange() dont have an impact on ggplot output. reorder(name,-median) orders name by -median (ascending order)

```{r}
w_plot <- ggplot(
  data = com_male,
  aes(x = reorder(name,-median_age), y = median_age)) +
  xlab(NULL) +
  ylab ("Age (in years)") +
  ggtitle ("Youngest Male Names")
```

__Finalizing the plot__

```{r}
w_plot + 
  geom_linerange(
    aes(ymin = q1_age, ymax = q3_age),
    color = "steelblue3", 
    size = 5, 
    alpha = 0.8
  ) + 
  geom_point(fill = "#ed3324", colour = "white", size = 3, shape = 21) +
  geom_point(aes(y=53,x=24),fill = "#ed3324", colour = "white", size = 3, shape = 21) +
  geom_text(aes(y=58,x=24),label="median",cex=3.8)+
  geom_text(aes(y=6,x=16),label="25th",cex=3.8)+
  geom_text(aes(y=22,x=16),label="75th percentile",cex=3.8)+
  geom_point(aes(y=8,x=16),shape=17)+
  geom_point(aes(y=17,x=16),shape=17)+
  coord_flip()

```

# Question 2 

List the top 10 years in which the name "Jackie was given to M and F babies most equally. 


```{r}
babynames2 <- babynames %>% 
  filter(name == "Jackie") %>% 
  group_by(year) %>% 
  summarise(
    total = sum(n),
    boys = sum(ifelse(sex == "M",n,0)),
    girls = total - boys) %>% 
  mutate(
    moeq = abs(.5 - boys/total)) %>% 
  arrange(moeq) %>%
  head(10)
  
babynames2 
```



## Two Table 

```{r}
library(Lahman)
library(dplyr)
```

Lahman data set has multiple tables of data including Master, Batter, AllStarFull and others. 

# Question 1 

Find top 20 HomeRun hitters 

```{r}

table <- Batting %>% 
  inner_join(Master, by = c("playerID" = "playerID"))
  

Top20HrHitters <- 
  newtable %>% 
  group_by(playerID) %>% 
  mutate(tHR = sum(HR), yearlast = max(yearID)) %>%
  filter(yearID == yearlast) %>% 
  select(nameFirst,nameLast,tHR) %>% 
  arrange(desc(tHR)) %>% 
  head(20) 
  
  
Top20HrHitters[,2:4]
```

# Question 2 

Summarize important statistics of Manny Ramirez's career and arrange by OPS.

```{r echo=FALSE}

#Create Manny by season

mannyBySeason <- Batting %>%
  filter(playerID == "ramirma02") %>%
  inner_join(Master, by = c("playerID"  = "playerID")) %>%
  group_by(yearID) %>%
  summarize(
    Age = max(yearID - birthYear), numTeams = n_distinct(teamID),
    BA = sum(H)/sum(AB), tH = sum(H), tHR = sum(HR), tRBI = sum(RBI),
    OBP = sum(H + BB + HBP) / sum(AB + BB + SF + HBP),
    SLG = sum(H + X2B + 2*X3B + 3*HR) / sum(AB)
  ) %>%
  mutate(OPS   = OBP + SLG) %>%
  arrange(desc(OPS))

mannyBySeason
```


# Tidying Data 


## Question 1 

Using spread and gather in conjuction. 

```{r}

stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)

stocks
```

```{r}
stocks %>%
  spread(year, return) %>%
  gather(`2015`:`2016`, key = "year", value = "return")
```