---
title: "Tidy Data"
author: "Charles Leahan"
date: "10/25/2019"
output: word_document
---

```{r}
library(tidyverse)
```

_Tidy Data_


Values are organised in two ways. Each value belongs to:

1. A variable, which contains all measurements of the same underlying attribute (like height, temperature, or duration) across units;

2. An observation (case), which contains all values measured on the same unit (like a person, day, or precinct) across attributes.


Three rules which make a dataset tidy:

* Each variable must have its own column
* Each observation must have its own row
* Each value must have its own cell

Data tidying refers to structuring datasets to facilitate analysis.



# Gather - Wide to Long

Collapses multiple columns into two columns:

1. Key -> former column names 
2. Value(n) -> former cells 


gather(data, name of the new key column, name of the new value column, names of numeric indexes of columns to collapse)


# Spread - Long to Wide

Generates multiple columns from two columns:

1. Each unique value in the key column becomes a column name 

2. Each value in the value column becomes a cell in the new columns 


spread(data frame to reshape, column to use for keys, column to use for values)

# Separate

Splits a column by a character string separator (like splitting dates for example)

separate(data,variable,c(sub variables), sep = "-")

# Unite

Unites columns in a single row

Does the the reverse 


## Example 1 

```{r}
#Source 

table1
```

```{r}
#Wide to long
table4a
gather(table4a,"year","n", 2:3)
```

```{r}
#Long to wide

table2
spread(table2,type,count)
```

# Missing Values

Explicitly - cell where its value should be instead contains NA

Implictly - does not appear in the dataset 


## Making Implicit Values Explicit

Spread():

In the new columns, NAs will be present where there wasn't a k/v pair

```{r}
stocks %>% 
  spread(key = year, value = return)
```

Complete():

Takes cart tuples of the fields provided, and returns them with NA if some combination
is not present in the data. 

```{r}
stocks %>% 
  complete(year,qtr)
```

Fill():

Carries forward last non-empty values.

```{r}
treatment %>% 
  fill(person)
```


# CASE STUDY

```{r}
head(who)
```

```{r}
#Gather

who1 <- who %>% 
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
who1
```

```{r}
#Zoom in the key 

who1 %>% 
  count(key)
```


```{r}
#Recode the key to make more clear
who2 <- who1 %>% 
  mutate(key = str_replace(key, "newrel", "new_rel"))
unique(who2$key)
```


```{r}

#Separate the key by "_" to extract more granular information

who3 <- who2 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")
who3
```


```{r}
#Separate the sexage key

who4 <- who3 %>% 
  separate(sexage, c("sex", "age"), sep = 1)
who4
```

```{r}
#Drop Redundant Columns (Variables)

who4 %>% 
  count(new)

who5 <- who4 %>% 
  select(-new, -iso2, -iso3)

head(who5)
```



str_replace(string, pattern, replacement)













```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

