---
title: "Supervised Learning (Intro)"
author: "Charles Leahan"
date: "12/2/2019"
output: word_document
---

```{r, echo = FALSE, warning=FALSE}
library(mdsr)
library(mosaicData)
```

Supervised Learning: 

Outcome variables are already measured and included in the data sets 

Find a function that accurately describes how different explanatory varaibles can be combined to explain the change of the response variable and to make a prediction.

$$y = f(x_{1},x_{2}, ..., x_{p}) + $\epsilon$$

Goals that motivate constructing a function: 

* Predict the output given an input (prediction)
* Determine which variables are useful inputs (association)
* Generate Hyp (causation)
* Understand how a system works 

Two Primary Problems: Regression (Predictive Modeling) and Classification 

# Regression 

Want to predict the ridership (volume) using other explanatory variables 

```{r}

#Plot relationship with superimposed loess line 

RailTrail %>% ggplot(aes(x = avgtemp, y = volume)) + 
  geom_point() +
  geom_smooth(method = 'loess')


```

The response, volume,is "supervising" the model fit

The scatter plot smoother (here "loess") is predicting volume. 

_Manually fitting a LOESS_

```{r}

#Fit a Loess model, and add the fitted values to the DF

RailTrail.fit <- RailTrail %>% 
  mutate(
    loess.fit = loess(volume ~ avgtemp, data = .)$fitted
  )

head(select(RailTrail.fit, volume, loess.fit, avgtemp))

```

```{r}
#Plot the fitted values, creating the same LOESS line as before

RailTrail.fit %>%
  ggplot() + 
  geom_point(aes(x = avgtemp, y = volume)) + 
  geom_line(aes(x = avgtemp, y = loess.fit), color = "blue",lwd = 1.5)
```


# Classification 

Use classifiers (explanatory variables) to predict grouping of ouput


## IRIS Data Example

Iris Data Set: 

Response: Species of IRIS Flower (setosa, virginica, versicolor)
Explanatory: Length and width of sepals and petals 

Can you classify these flowers using the measurement Petal.Length into one of the species?  A crude answer is as follows.

Petal.Length < 2 : setosa
2 < Petal.Length <5 : versicolor
Petal.Length > 5 : virginica


```{r}

#Crude Classification by the above rules  

iris.crude <- iris %>% mutate(true.Species = Species) %>%
  mutate(
    predicted.Species = 
      ifelse(Petal.Length < 2, "setosa", 
             ifelse(Petal.Length < 5, "versicolor", 
                    "virginica"))
    ) 

set.seed(125)

sample_n(select(iris.crude, Petal.Length, true.Species, predicted.Species),10)

```

```{r}
#Plot the Crude Classification 

iris.crude %>% 
  ggplot(aes(x = Petal.Length, y = true.Species,  
             color = predicted.Species)) + 
  geom_point()

```

Using the previous classification as well as Petal.Width

```{r}
#Crude classification with multiple variables

iris.crude2 <- iris %>% mutate(true.Species = Species) %>%
  mutate(
    predicted.Species = 
      ifelse(Petal.Length < 2 & Petal.Width < 0.75, "setosa", 
             ifelse(Petal.Length < 5 & Petal.Width < 1.75, "versicolor", 
                    "virginica"))
    ) 

#Plot the Classification

iris.crude2 %>% 
  ggplot(aes(x = Petal.Length, y = Petal.Width, 
             color = true.Species, shape = predicted.Species)) + geom_point()
```

## Obama-Clinton example

Use the Obama-Clinton divide dataset 

```{r, echo = FALSE}
primary <- read.csv('C:/Users/CPL17/py_all/data_sets/primaries.csv')
```


__Exploration 1__ 

```{r}
#Exploration 

primary %>% ggplot(aes(x = region, fill = winner)) + 
  geom_bar(position = "fill")  
```

__Exploration 2__

```{r}
#Exploration: See if Bush04 is a good indication of winner, observe conditional distribution of Bush04

primary <- primary %>% filter(!is.na(winner))
primary %>% 
  ggplot() + 
  geom_boxplot(aes(x = winner, y = Bush04))+
  coord_flip()
```

__Exploration 3__

```{r}

#Exploration: See if Black and pop are good indicators

primary %>% ggplot(aes(x = black06, y = popUnder30_00, color = winner)) + geom_point()


```


```{r}
#Transforation: Re-scale black to account for difference in size of counties

primary <- 
  primary %>% 
  mutate(AFprop = black06 / pop06 * 100) %>% 
  filter(!is.na(AFprop))
```


```{r}

#Plot with scale log(10) to deal with small AFprop values 

primary %>% ggplot(aes(x = AFprop, y = popUnder30_00, 
                       color = winner)) + 
  geom_point() +
  scale_x_log10() 

```





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


