---
title: 'Lecture 17 :  Statistical Modeling (2)'
author: "Charles Leahan"
date: "10/23/2019"
output: word_document
---



# Categorical Explanatory Variables 


## Example: Railtrail

Exaplain volume by the binary variable weekday (= 1 if weekday)

```{r}
#Distribution of weekday

library(mosaicData)

RailTrail %>% 
  group_by(weekday) %>%
  summarise(mean_volume = mean(volume))
```

```{r}
#Box plot to visualize relationship

RailTrail %>% 
  ggplot(aes(x = weekday, y= volume)) + 
  geom_boxplot()+
  coord_flip()
```


```{r}
#Fit the SLR 

mod2 <- lm(volume ~ weekday, data = RailTrail)
mod2$coefficients
```

Note:

* lm picks the alphabetically lowest value of the categorical predictor as the reference group
* beta0 is the mean volume given x = 0
* 80 fewer in volume on weekdays





# Multiple Regression 

```{r}
#Fit mlr model

mod3 <- lm(volume ~ hightemp + weekday + hightemp*weekday,  data = RailTrail)
summary(mod3)
```

```{r}
#Plot the model

plotModel(mod3,system = "ggplot2")
```


```{r}
mod4 <- lm(volume ~ hightemp + cloudcover + weekday,  data = RailTrail)
summary(mod4)
```

```{r}

# will use all variables other than volume 

mod5 <- lm(volume ~ .,  data = RailTrail)
summary(mod5)
```

Note: 

* NA variables arent estimated due to very high collinearity with other variables. Can be dealt with using regularized or sparse regression. 



# Model Selection 

Note:

* install "leads" package to get more criterea for model selection

```{r}
library(MASS)
mod6 <- lm(volume ~ hightemp + lowtemp + spring + summer + weekday + cloudcover + precip,data=RailTrail)
step <- stepAIC(mod6,direction = 'both')
```


# Non-Parametric Regression - Scatterplot smoothing 

```{r}
RTsim <- RailTrail %>% transmute(x = hightemp, y = volume)
scat_ex <- RTsim %>% ggplot(aes(x, y)) + geom_point()

```

```{r}
scat_ex + geom_smooth(span = 10)
scat_ex + geom_smooth(span = 1/5)
```

Note:

*The second method treats noise as signal

# Data Transformation: Numeric to Categorical 

```{r}
SF %>% 
  mutate(long_delay = arr_delay > 60) %>%
  ggplot(aes(x = hour,  fill= long_delay)) + 
  geom_bar(position = "fill") +
  xlab("Scheduled hour of departure") + 
  ylab("Proportion of long delay") 
```

```{r}
SF_longdelay <- SF %>% mutate(long_delay = arr_delay > 60) 
SF_longdelay %>% ggplot(aes(x = hour,  y = long_delay)) +geom_point()
```























```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=4, fig.height=2)
```


