---
title: ''
author: "Charles Leahan"
date: "11/11/2020"
output: word_document
---

# Adding Context 

```{r}
g <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE)
```


```{r}
#Use labs() titles,captions

g + labs(
       title = "Fuel efficiency generally decreases with engine size",
       subtitle = "Two seaters (sports cars) are an exception because of their light weight",
       caption = "Data from fueleconomy.gov")
```

```{r}
#Use labs to change x,y axis labels

g + labs(x = "Engine displacement (L)",y = "Highway fuel economy (mpg)",colour = "Car type")

```

# Scales 

```{r}
#Compare price and log price

par(mfrow = c(1,2))
ggplot(diamonds, aes(carat,price)) + geom_bin2d()
ggplot(diamonds, aes(carat,log10(price))) + geom_bin2d()
```

```{r}
#Adjust Scale 

ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() + 
  scale_x_log10() + 
  scale_y_log10()
```

# Change colors 

```{r}
#Use scale_colour_brewer to set new color pallete

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  scale_colour_brewer(palette = "Set1")
```

# Creating proportion and frequency charts

```{r}
#Plot cout of each type of cut 

ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```

## Method 1 

```{r}
#Proportion

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1)) #group = 1 -> group by prop
```

Note: 

..stat.. performs the statistical function/aggregation on the present field (x)


## Method 2 

```{r}
#Proportion (Manually)

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..count../sum(..count..)))+
  ylab("Proportion")
```

# Method 3 

```{r}
library(tibble)
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)
```

Note: 

tribble() creates tibbles using an easier to read row-row format. 

```{r}
#Plot 
ggplot(data = demo) +
  geom_bar(mapping = aes(x = reorder(cut,freq), y = freq), stat = 'identity')
```

Note: 

reorder(x,y) reoders x by y 

stat = 'count' is the default -> counts the # of cases at each x position, cannot be used with y aesthetic 

stat = 'identity -> leaves the data as is

geom_col -> does the above graph and uses stat_identity()

## Changing the color 

```{r}
#Change the color of the bars

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut,fill = cut)) 
```

```{r}
#Change transparancy within the bars

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut,color = cut),alpha=0.5)
```

## Adding a variable

```{r}
#Using clarity to fill 

g <- ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) 
g + geom_bar() 
```

## Position Arguments 

```{r}
g + geom_bar(position = "fill")  #Shows conditional proportions
g + geom_bar(position = "dodge") #Shows distribution
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


